---
title: "Report for assignment 1"
author: "Wei Jiang"
date: "Saturday, February 07, 2015"
output: html_document
---

This is the report for assignment 1, for Reproducible Research.

### Loading and preprocessing the data
First, download and extract the dataset [Activity monitoring data](https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip), we got activity.csv. Then load the data and check the summary.

```{r}
## setwd("d:/R/RStudio/coursera/5.Reproducible_research")
unzip("./activity.zip")
rdata <- read.csv("activity.csv",header=TRUE,as.is=TRUE)
rdata$date <- as.Date(strptime(rdata$date, format="%Y-%m-%d"))
rdata$date.f <- as.factor(rdata$date)
head(rdata)
```

### What is the mean total number of steps taken per day?

```{r}
x <- data.frame(levels(rdata$date.f),tapply(rdata$steps, rdata$date.f, sum))
colnames(x) <- c("date","total.steps")
rownames(x) <- c(1:dim(x)[1])
summary(x)
xx <- x[which(!is.na(x$total.steps)),]
xx$total.steps <- as.numeric(xx$total.steps)

stepsMean <- round(mean(xx$total.steps))
stepsMedian <- median(xx$total.steps)

library(lattice)
histogram(~total.steps, data=xx,
    xlab="total steps per day",
    breaks = 20,
    panel=function(x,...) {
        panel.histogram(x,...)
        panel.abline(v=stepsMean,col="red",lty=3,lwd=3)
        panel.abline(v=stepsMedian,col="black",lty=2,lwd=2)
    }
)
```

The mean of total numbers of steps (per day) is `r stepsMean`, median is `r stepsMedian`.

### What is the average daily activity pattern?
``` {r fig.width=10}
library(ggplot2)
rdata$interval.f <- as.factor(rdata$interval)
x <- rdata[which(!is.na(rdata$steps)),]
xx <- data.frame(levels(x$interval.f),tapply(x$steps, x$interval.f, mean))
colnames(xx) <- c("interval.f","average.steps")
xx$interval <- as.numeric(rownames(xx))

qplot(interval, average.steps, data=xx, geom=c("line","smooth"), method="lm")

result <- as.character(xx[which(xx$average.steps==max(xx$average.steps)),][[1]])
```

In 5-minute interval, across all the days, on average, the interval with maximum number of steps is `r result`.

### Inputing missing values

By checking the summary, only "steps" has missing values, and there are 2304 NA's:
```{r}
summary(rdata[,1:3])
```

```{r}
## xx is the variable contains average steps for each interval
x <- rdata
for (i in 1:length(rdata$steps)) {
    ## when steps is missing
    if(is.na(x[i,1])) {
        ## find the same row's internal value
        intervalValue <- x[i,3]
        ## get the average steps from xx based on the same internal value
        x[i,1] <- xx[which(xx$interval==intervalValue),][[2]]
    }
}
```
Now the new dataset looks like this: 
```{r}
summary(x[,1:3])
```

```{r}
stepsPerDay <- data.frame(levels(rdata$date.f),tapply(x$steps, x$date.f, sum))
colnames(stepsPerDay) <- c("date","total.steps")
stepsMean <- round(mean(stepsPerDay$total.steps))
stepsMedian <- median(stepsPerDay$total.steps)

library(lattice)
histogram(~total.steps, data=stepsPerDay,
    xlab="total steps per day",
    breaks = 20,
    panel=function(x,...) {
        panel.histogram(x,...)
        panel.abline(v=stepsMean,col="red",lty=3,lwd=3)
        panel.abline(v=stepsMedian,col="black",lty=2,lwd=2)
    }
)
```

After filling in the missing values, the mean of total numbers of steps (per day) is `r stepsMean`, median is `r stepsMedian`. 

Compared with first part of this report, the mean value was not changed, but the median was changed. It is a little obvious, since we filled in the missing values with average steps, so those values was originally not counted, but now devoted to the mean with a average value, so their devoting is nothing.

However, since those newly added values, we now got a large amount of points to elect the median, so the median value was different from first part.


### Are there differences in activity patterns between weekdays and weekends?

```{r}
x <- rdata[which(!is.na(rdata$steps)),]
library(chron)
x$weekend <- is.weekend(x$date)


##xx_weekend <- data.frame(levels(x$interval.f),tapply(x$steps, x$interval.f, mean))
colnames(xx) <- c("interval.f","average.steps")
xx$interval <- as.numeric(rownames(xx))

x_weekend <- x[which(x$weekend),]
x_weekday <- x[which(!x$weekend),]

xx_weekend <- data.frame(levels(x_weekend$interval.f),tapply(x_weekend$steps, x_weekend$interval.f, mean))
colnames(xx_weekend) <- c("interval","average.steps")
xx_weekend$weekend <- "weekend"
xx_weekday <- data.frame(levels(x_weekday$interval.f),tapply(x_weekday$steps, x_weekday$interval.f, mean))
colnames(xx_weekday) <- c("interval","average.steps")
xx_weekday$weekend <- "week day"
xx <-rbind(xx_weekday,xx_weekend)
xx$weekend <-as.factor(xx$weekend)
xx$interval <- as.numeric(xx$interval)

g <- ggplot(xx, aes(interval, average.steps))
g + geom_line(aes(color=weekend))  + geom_smooth(method="lm")  +facet_grid(.~weekend) + labs(x="Intervals (every 5 minutes)") + labs(y="Average steps")

```

Yes, on week day, people take more activity at end of day, while during weekend, people don't take that much activities.
